Design and Implementation of Probabilistic Programming Language Anglican

### 2016-08-08 paper/paper.tex
update

Operational semantics of Anglican queries is different from that
of Clojure code, therefore queries must be called through
inference algorithms, rather than `directly'.  The
\texttt{anglican.inference} namespace declares the polymorphic
function \texttt{infer} using Clojure's multimethod mechanism.
This function accepts an Anglican query and
returns a lazy sequence of weighted samples from the
distribution defined by the query.  When inference
is performed on an Anglican query, the query is run by a
particular inference algorithm. Inference algorithms must
provide an implementation for \texttt{infer}, as well as
override the polymorphic function \texttt{checkpoint} (defined
as a multimethod) so as to handle \texttt{sample} and
\texttt{observe} in an algorithm-specific manner and
to construct an appropriate result on the 
termination of a probabilistic program.

@@

An inference algorithm is an implementation of \texttt{infer}
multimethod from \texttt{anglican.inference} namespace. The
multimethod accepts an algorithm identifier, a query --- the
probabilistic program in which to perform the inference, an
initial value for the query, and optional algorithm parameters.

\subsection{The \texttt{infer} multimethod}

The sole purpose of the algorithm identifier is method dispatch
--- conventionally, the identifier is a keyword related to the
algorithm name (\texttt{:lmh} for Lightweight
Metropolis-Hasting, \texttt{:pcascade} for Particle Cascade
etc.). The second parameter is a query as defined by
\texttt{query} or \texttt{defquery} forms. If needed, the query
can be defined anonymously, right in the argument list of a call
to \texttt{infer}:
\begin{lstlisting}[style=default]
(let [x 1]
  (infer :pgibbs (query (predict x)) nil))
\end{lstlisting}
A query is executed by calling the initial continuation of the
query, which accepts a value and a state. The state is supplied
by the inference algorithm, while the value is provided as a
parameter of \texttt{infer}. A query does not have to have any
parameters, in which case the value can be simply \texttt{nil}.
When a query is defined with a binding for the initial value,
the value becomes available inside the query. A query may
accept multiple parameters as components of a Clojure vector,
in which case the initial value is given as a vector and its
components get bound to the corresponding parameters of the
query (via destructuring).
%The value is destructured via pattern match in case of a structured binding. 
For example,
\begin{lstlisting}[style=default]
(defquery my-query [mean sd]
  (predict (sample (normal mean sd))))

>>>>

Operational semantics of Anglican queries is different from that
of Clojure code, therefore queries must be called through
inference algorithms, rather than `directly'.  The
\texttt{anglican.inference} namespace declares the (ad-hoc) polymorphic
function \texttt{infer} using Clojure's multimethod mechanism.
This function accepts an Anglican query and
returns a lazy sequence of weighted samples from the
distribution defined by the query.  When inference
is performed on an Anglican query, the query is run by a
particular inference algorithm. Inference algorithms must
provide an implementation for \texttt{infer}, as well as
override the polymorphic function \texttt{checkpoint} (defined
as a multimethod) so as to handle \texttt{sample} and
\texttt{observe} in an algorithm-specific manner and
to construct an appropriate result on the 
termination of a probabilistic program.

@@

An inference algorithm is an implementation of the (ad-hoc) 
polymorphic function \texttt{infer} declared as Clojure's multimethod
in the \texttt{anglican.inference} namespace. The
function accepts an algorithm identifier, a query --- the
probabilistic program in which to perform the inference, an
initial value for the query, and optional algorithm parameters.

\subsection{The \texttt{infer} function}

The sole purpose of the algorithm identifier of \texttt{infer} is to invoke
an appropriate overloading or implementation of the function --- conventionally, 
the identifier is a keyword related to the
algorithm name (\texttt{:lmh} for Lightweight
Metropolis-Hasting, \texttt{:pcascade} for Particle Cascade
etc.). The second parameter is a query as defined by
the \texttt{defquery} form or its anonymous version \texttt{query}. For instance,
the following Clojure code invokes \texttt{infer} on an Anglican query
defined anonymously via the \texttt{query} form:
%If needed, the query
%can be defined anonymously, right in the argument list of a call
%to \texttt{infer}:
\begin{lstlisting}[style=default]
(let [x 1]
  (infer :pgibbs (query (predict x)) nil))
\end{lstlisting}
A query is executed by calling the initial continuation of the
query, which accepts a value and a state. The state is supplied
by the inference algorithm, while the value is provided as a
parameter of \texttt{infer}. A query does not have to have any
parameters, in which case the value can be simply \texttt{nil}.
When a query is defined with a binding for the initial value,
the value becomes available inside the query. A query may
accept multiple parameters using Clojure's structured binding.
For instance, it may take multiple parameters as components 
of an input vector. In this case, the initial value is given
as a structured value, such as a vector, and the components
of this value gets pattern matched to the corresponding
parameters of the query via the destructuring mechanism of Clojure:
%The value is destructured via pattern match in case of a structured binding. 
For example,
\begin{lstlisting}[style=default]
(defquery my-query [mean sd]
  (predict (sample (normal mean sd))))
